```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)                        
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
library(car)
library(microViz)
```

```{r}
getwd()
```

```{r}
###Input 16S 
otus.bac <- read.table("./Baceria/Data/CrustTemp216SasvUN3.asv_table_fix.txt", header = TRUE, sep = "\t", row.names = 1)
colnames(otus.bac) <- gsub("X", "", colnames(otus.bac))  # Remove "X" from all column names
otumat16s <- as.matrix(otus.bac)
OTU16s <- otu_table(otumat16s, taxa_are_rows = TRUE)
```

```{r}
taxmat16s <- read.table("./Baceria/Data/bac_tax_fix.csv", header=T,sep=",",row.names=1)
taxmat16s <- as(as.matrix(taxmat16s),"matrix")
TAX16s <- tax_table(taxmat16s)
```

```{r}
meta16s <- read.table("./Baceria/Data/CrustTemp216SasvUN3.mapping_file_fix.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)

sampleData16s <- sample_data(meta16s)
```

```{r}
bac.tree <- read_tree("./Baceria/Data/tree.nwk")
```

```{r}
physeq16s <- phyloseq(OTU16s,TAX16s,sampleData16s, bac.tree)
physeq16s
```
```{r}

physeq16s <- subset_taxa(physeq16s, Kingdom == "Bacteria")
physeq16s
physeq16s<- subset_taxa(physeq16s, Family != "Mitochondria")
physeq16s
#physeq16s <- subset_taxa(physeq16s, Species != "NA")
#physeq16s <- subset_taxa(physeq16s, Species != "uncultured bacterium")
#physeq16s <- subset_taxa(physeq16s, Species != "metagenome")

physeq16S.prune = prune_taxa(taxa_sums(physeq16s) > 1, physeq16s)
physeq16S.prune
```

### Rarefraction curve

Check read counts: any samples that have very low reads should be removed.
[Ref](http://evomics.org/wp-content/uploads/2016/01/phyloseq-Lab-01-Answers.html)

```{r}
readcount = data.table(as(sample_data(physeq16S.prune), "data.frame"),
                 TotalReads = sample_sums(physeq16S.prune), 
                 keep.rownames = TRUE)
setnames(readcount, "rn", "SampleID")

#For plotting, use command below.
SeqDepth = ggplot(readcount, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
```

TotalReads of all the samples can be in this table (select only SampleID and TotalReads columns).
In order to check samples with low number of reads, "order()" can be used to sort "TotalReads" column.
In this dataset, N55.Rhizo has very low number of reads, so will will filter this sample out using the next minimum number of reads.
```{r}
readcount = readcount[order(readcount$TotalReads), c("SampleID", "TotalReads")]
```

```{r}
as.data.frame(readcount)
```

Extract OTU table, transpose and convert to data frame
Extract OTU table from phyloseq object as data frame and use vegan to calculate rarefaction curve
```{r}
otu.rare <- otu_table(physeq16S.prune)
otu.rare <- as.data.frame(t(otu.rare))
sample_names <- rownames(otu.rare)

#vegan rarecurve 
png("D:/Biocrust_Reawakening/Github/Bacteria(16S)/Figures/Rarefraction16S.png", units="in", width = 8.5, height = 6, res = 600 )
otu.rarecurve <- rarecurve(otu.rare, step = 500, sample = 55000, label = T)
#dev.off()
```

```{r}
otu.rare <- otu_table(physeq16S.prune)
otu.rare <- as.data.frame(t(otu.rare))
sample_names <- rownames(otu.rare)

#vegan rarecurve 
png("D:/Biocrust_Reawakening/Github/Fungi(ITS)/Figures/Rarefraction.png", units="in", width = 9.5, height = 7, res = 600 )
otu.rarecurve <- rarecurve(otu.rare, step = 100, sample = 50000, label = T)
dev.off()
```


### Taxonomic composition


```{r}
physeq.LAC <- subset_samples(physeq16S.prune, Crusttype=="LAC")
physeq.LAC

physeq.CLC <- subset_samples(physeq16S.prune, Crusttype=="CLC")
physeq.CLC
```

```{r}
glom.bac.LAC <- speedyseq::tax_glom(physeq.LAC,taxrank = "Phylum")
glom.bac.LAC

glom.bac.CLC <- speedyseq::tax_glom(physeq.CLC,taxrank = "Phylum")
glom.bac.CLC
```


```{r}
head(tax_table(glom.bac.LAC))
```

```{r}
bac.abund.LAC <- merge_samples(glom.bac.LAC, "Incubation")
sample_data(glom.bac.LAC)$Incubation <- factor(sample_names(glom.bac.LAC))
bac.abund.LAC = transform_sample_counts(bac.abund.LAC, function(x) x / sum(x))
bac.abund.LAC

bac.abund.CLC <- merge_samples(glom.bac.CLC, "Incubation")
sample_data(glom.bac.CLC)$Incubation <- factor(sample_names(glom.bac.CLC))
bac.abund.CLC = transform_sample_counts(bac.abund.CLC, function(x) x / sum(x))
bac.abund.CLC
```
```{r}
data_glom.bac.LAC <- speedyseq::psmelt(bac.abund.LAC)
data_glom.bac.LAC$Phylum <- as.character(data_glom.bac.LAC$Phylum)

data_glom.bac.CLC <- speedyseq::psmelt(bac.abund.CLC)
data_glom.bac.CLC$Phylum <- as.character(data_glom.bac.CLC$Phylum)
```

```{r}
data_glom.bac.LAC$Phylum[data_glom.bac.LAC$Abundance < 0.005] <- "others"
data_glom.bac.CLC$Phylum[data_glom.bac.CLC$Abundance < 0.005] <- "others"
```

```{r}
# Extract percentage data
percentage_data <- data_glom.bac.CLC %>%
  group_by(Incubation, Phylum) %>%
  summarize(Percentage = sum(Abundance)) %>%
  group_by(Incubation) %>%
  mutate(Percentage = Percentage / sum(Percentage) * 100) %>%
  gather(key = "Phylum", value = "Percentage", -Incubation)

# View the percentage data
print(percentage_data)
```

```{r}
write.csv(percentage_data, file = "D:/Biocrust_Reawakening/New_data/Bacteria/CLC_Genus_percentage_data.csv", row.names = FALSE)
```


```{r}
# Merge the two datasets for consistent color mapping
data_glom_combined <- bind_rows(
  mutate(data_glom.bac.LAC, Group = "LAC"),
  mutate(data_glom.bac.CLC, Group = "CLC")
)

# Ensure consistent levels for the "Genus" factor
#data_glom_combined$Genus <- factor(data_glom_combined$Genus, levels = names(cbbPalette))
```

```{r}
Count = length(unique(data_glom_combined$Phylum))
Count
```

```{r}
unique((data_glom_combined$Phylum))
```

```{r}
data_glom_combined$Genus <- factor(data_glom_combined$Genus, levels = c(
"unclassified_o_Oxyphotobacteria Incertae Sedis",
"unclassified_o_Nostocales" ,
"unclassified_f_Coleofasciculaceae" ,
"unclassified_f_Cyanobacteriaceae"  ,
"unclassified_f_Nostocaceae" ,
"Tychonema CCAP 1459-11B" ,
"Symplocastrum CPER-KK1" ,
"Microcoleus SAG 1449-1a" ,
"Microcoleus PCC-7113" ,
"Wilmottia Ant-Ph58" ,
"Trichocoleus SAG 26.92"  ,
"Mastigocladopsis PCC-10914"   ,
"Microcoleus Es-Yyy1400" ,
"Nostoc PCC-73102",
"Scytonema UTEX 2349" ,
"CENA359" ,
"Aliterella CENA595",

"unclassified_f_AKIW781",
"unclassified_f_Ardenticatenaceae",
"unclassified_f_Roseiflexaceae",

"unclassified_f_Burkholderiaceae"   ,
"unclassified_f_Caulobacteraceae" ,
"unclassified_f_Sphingomonadaceae"    ,
"unclassified_f_Archangiaceae" ,
"unclassified_f_Myxococcaceae"  ,
"unclassified_f_Beijerinckiaceae" ,
"unclassified_f_0319-6G20",
"unclassified_f_Acetobacteraceae",
"unclassified_f_Polyangiaceae",
"Microvirga" ,
"Azospirillum",
"Rubellimicrobium" ,
"Massilia" ,
"Noviherbaspirillum"  ,
"Ellin6055" , 
"Craurococcus" ,
"Magnetospirillum sp. enrichment culture clone Van25" ,
"MN 122.2a" ,
"Nitrospirillum" ,
"Paucimonas" ,
"Ramlibacter" ,
"Sphingoaurantiacus" ,
"Telmatospirillum" ,

"unclassified_f_Micromonosporaceae" ,
"Geodermatophilus" ,
"Rubrobacter"  ,
"Arthrobacter",
"Pseudonocardia" ,

"unclassified_f_Spirosomaceae" ,
"Cnuella" ,
"Adhaeribacter" ,
"Hymenobacter" ,
"Rhodocytophaga" ,

"unclassified_f_Family XVIII" ,
"unclassified_f_Paenibacillaceae" ,
"unclassified_f_Planococcaceae" ,
"Clostridium sensu stricto 10",
"Tumebacillus" ,
"Paenibacillus" ,
"Bacillus" ,
"Clostridium sensu stricto 8", 
"Clostridium sensu stricto 1" ,
"Brevibacillus" ,
"Azohydromonas",
"Fonticella" ,

"unclassified_f_Gemmataceae",
"unclassified_f_Longimicrobiaceae" ,
"Gemmatirosa" ,
"Longimicrobium" ,

"unclassified_f_Blastocatellaceae",
"Blastocatella" ,

"unclassified_f_Fimbriimonadaceae",
"others"
))
```

```{r}

cbbPalette <- c(
  # --- CYANOBACTERIA (Greens) ---
  "unclassified_o_Oxyphotobacteria Incertae Sedis" = "#458B00", # chartreuse4
  "unclassified_o_Nostocales"                      = "#66CD00", # chartreuse3
  "unclassified_f_Coleofasciculaceae"              = "#76EE00", # chartreuse2
  "unclassified_f_Cyanobacteriaceae"               = "#A2CD5A", # darkolivegreen3
  "unclassified_f_Nostocaceae"                    = "#C0FF3E", # chartreuse1
  "Tychonema CCAP 1459-11B"                        = "#556B2F", # darkolivegreen
  "Symplocastrum CPER-KK1"                         = "#6E8B3D", # darkolivegreen4
  "Microcoleus SAG 1449-1a"                        = "#8FBC8F", # darkseagreen
  "Microcoleus PCC-7113"                           = "#228B22", # forestgreen
  "Wilmottia Ant-Ph58"                             = "#006400", # darkgreen
  "Trichocoleus SAG 26.92"                         = "#2E8B57", # seagreen
  "Mastigocladopsis PCC-10914"                     = "#3CB371", # mediumseagreen
  "Microcoleus Es-Yyy1400"                         = "#00FF7F", # springgreen
  "Nostoc PCC-73102"                               = "#00C957", # emerald green
  "Scytonema UTEX 2349"                            = "#90EE90", # lightgreen
  "CENA359"                                        = "#00FA9A", # mediumspringgreen
  "Aliterella CENA595"                             = "#BDFCC9",

  # --- CHLOROFLEXI (Oranges/Siennas) ---
  "unclassified_f_AKIW781"                         = "#CD661D", # sienna3
  "unclassified_f_Ardenticatenaceae"               = "#E69F00", # orange
  "unclassified_f_Roseiflexaceae"                  = "#FF8C00", # darkorange

  # --- PROTEOBACTERIA (Blues) ---
  "unclassified_f_Burkholderiaceae"                = "#104E8B", # dodgerblue4
  "unclassified_f_Caulobacteraceae"                = "#1E90FF", # dodgerblue
  "unclassified_f_Sphingomonadaceae"               = "#63B8FF", # dodgerblue1
  "unclassified_f_Archangiaceae"                   = "#00BFFF", # deepskyblue
  "unclassified_f_Myxococcaceae"                   = "#00008B", # darkblue
  "unclassified_f_Beijerinckiaceae"                = "#4682B4", # steelblue
  "unclassified_f_0319-6G20"                       = "#5CACEE", # steelblue2
  "unclassified_f_Acetobacteraceae"                = "#009ACD", # deepskyblue3
  "unclassified_f_Polyangiaceae"                   = "#00CED1", # darkturquoise
  "Microvirga"                                     = "#add8e6", # lightblue
  "Azospirillum"                                   = "#87CEFA", # lightskyblue
  "Rubellimicrobium"                               = "#B0E2FF", # lightblue1
  "Massilia"                                       = "#00441B", # deep teal
  "Noviherbaspirillum"                             = "#2B8CBE", 
  "Ellin6055"                                      = "#A6BDDB",
  "Craurococcus"                                   = "#36648B",
  "Magnetospirillum sp. enrichment culture clone Van25" = "#5F9EA0", # cadetblue
  "MN 122.2a"                                      = "#00688B", # deepskyblue4
  "Nitrospirillum"                                 = "#708090", # slate gray
  "Paucimonas"                                     = "#4F94CD", # steelblue3
  "Ramlibacter"                                    = "#B0C4DE", # lightsteelblue
  "Sphingoaurantiacus"                             = "#6495ED", # cornflowerblue
  "Telmatospirillum"                               = "#0000CD", # mediumblue

  # --- ACTINOBACTERIA (Creams/Purples) ---
  "unclassified_f_Micromonosporaceae"              = "#FFFACD", # lemonchiffon
  "Geodermatophilus"                               = "#EEDC82", # silk
  "Rubrobacter"                                    = "#CDC673", # darkkhaki
  "Arthrobacter"                                   = "#FFF8DC", # cornsilk
  "Pseudonocardia"                                 = "#FFEFD5", # papayawhip

  # --- BACTEROIDETES (Tan/Browns) ---
  "unclassified_f_Spirosomaceae"                   = "#DEB887", # burlywood
  "Cnuella"                                        = "#D2B48C", # tan
  "Adhaeribacter"                                  = "#BC8F8F", # rosybrown
  "Hymenobacter"                                   = "#8B7355", # burlywood4
  "Rhodocytophaga"                                 = "#CDAA7D", # burlywood3

  # --- FIRMICUTES (Chocolates/Reds) ---
  "unclassified_f_Family XVIII"                    = "#D2691E", # chocolate
  "unclassified_f_Paenibacillaceae"                = "#8B4513", # saddlebrown
  "unclassified_f_Planococcaceae"                  = "#A0522D", # sienna
  "Clostridium sensu stricto 10"                   = "#FF4500", # orangered
  "Tumebacillus"                                   = "#CD3333", # brown3
  "Paenibacillus"                                  = "#8B2323", # brown4
  "Bacillus"                                       = "#BC1717", 
  "Clostridium sensu stricto 8"                    = "#800000", # maroon
  "Clostridium sensu stricto 1"                    = "#B22222", # firebrick
  "Brevibacillus"                                  = "#A52A2A", # brown
  "Azohydromonas"                                  = "#E9967A", # darksalmon
  "Fonticella"                                     = "#FA8072", # salmon

  # --- GEMMATIMONADETES (Golds) ---
  "unclassified_f_Gemmataceae"                     = "#DAA520", # goldenrod
  "unclassified_f_Longimicrobiaceae"               = "#B8860B", # darkgoldenrod
  "Gemmatirosa"                                    = "#FFD700", # gold
  "Longimicrobium"                                 = "#EEB422", # goldenrod2

  # --- ACIDOBACTERIA/OTHERS (Various) ---
  "unclassified_f_Blastocatellaceae"               = "#FFD300",
  "Blastocatella"                                  = "#CDAD00", # gold3
  "unclassified_f_Fimbriimonadaceae"               = "#9ACD32", # yellowgreen
  "others"                                         = "#868280"
)

```

```{r}
data_glom_combined$Phylum <- factor(data_glom_combined$Phylum, levels = c(
"Cyanobacteria",
"Proteobacteria",
"Firmicutes" ,
"Actinobacteria",
"Bacteroidetes" ,     
"Gemmatimonadetes" ,
"Acidobacteria",
"FBP" ,
"Chloroflexi",
"Planctomycetes" ,
"Verrucomicrobia",
"Deinococcus-Thermus",
"Armatimonadetes" ,
"Nitrospirae" ,
"Patescibacteria"  ,
"Fibrobacteres" ,
"WPS-2",
"others" 
))

```
```{r}
cbbPalette <- c(
  "Cyanobacteria"       = "chartreuse3",
  "Proteobacteria"      = "skyblue3",
  "Firmicutes"          = "chocolate3",
  "Actinobacteria"      = "antiquewhite2",
  "Bacteroidetes"       = "burlywood1",
  "Gemmatimonadetes"    = "darkgoldenrod4",
  "Acidobacteria"       = "gold3",
  "FBP"                 = "plum3",
  "Chloroflexi"         = "sienna2",
  "Planctomycetes"      = "darkgoldenrod1",
  "Verrucomicrobia"     = "orchid4",
  "Deinococcus-Thermus" = "mediumpurple1",
  "Armatimonadetes"     = "gold4",
  "Nitrospirae"         = "firebrick",
  "Patescibacteria"     = "lightseagreen",
  "Fibrobacteres"       = "darkolivegreen4",
  "WPS-2"               = "tomato3",
  "others"              = "#868280"
)
```

```{r}
# Ensure consistent levels for the "Genus" factor
data_glom_combined$Genus <- factor(data_glom_combined$Genus, levels = names(cbbPalette))
```

```{r}
data_glom_combined$Incubation = factor(data_glom_combined$Incubation, levels = c("0.05","1","24","48","96"))
data_glom_combined$Group <- factor(data_glom_combined$Group, levels = c("LAC", "CLC"))
```

```{r}
data_glom_combined <- data_glom_combined[!grepl("others", data_glom_combined$Genus), ]
```

```{r}
# Define your standard sizes
size_title <- 14
size_axis <- 14
size_label <- 10
main_font <- "serif" # Time news roman
```

```{r}
facet_wrap(
  ~Group,
  labeller = labeller(Group = c(
    LAC = "Light cyanobacteria crust",
    CLC = "Cyanolichen crust"
  ))
)
```
```{r}
# Italicize only the genus (first word)
genus_labels <- function(x) {
  sapply(x, function(i) {

    # Leave unclassified or code-only names as plain text
    if (grepl("^unclassified", i) || grepl("^[A-Z0-9-]+$", i)) {
      return(paste0("'", i, "'"))   # <- quote the whole label
    }

    parts <- strsplit(i, " ", fixed = TRUE)[[1]]

    if (length(parts) == 1) {
      paste0("italic('", parts[1], "')")
    } else {
      paste0(
        "italic('", parts[1], "')~'",
        paste(parts[-1], collapse = " "),
        "'"
      )
    }
  })
}
```

```{r fig.height=5, fig.width=8, fig.align="center"}
jpeg("./Baceria/Figures/Taxcom_Phylum_Bac_0.05.jpeg", units="in", width = 7.5, height = 6, res = 600 )

# Create the combined plot
plot_Fun_Phylum <- ggplot(
  data_glom_combined,
  aes(x = Incubation, y = Abundance, fill = Phylum)
) +
  geom_bar(stat = "identity", position = "fill") +
  scale_x_discrete(labels = c("0.05H", "1H", "24H", "48H", "96H")) +
  scale_fill_manual(values = cbbPalette) +
                    #labels = function(x) parse(text = genus_labels(x))) 
                    
  facet_wrap(
    ~Group,
    labeller = labeller(Group = c(
      LAC = "Light cyanobacterial crust",
      CLC = "Cyanolichen crust"
    ))
  ) +
  ggtitle("Bacterial Taxonomic Composition") +
  theme_bw() +
  theme_pubr(border = TRUE) +
  theme(
    text = element_text(family = main_font),
    plot.title = element_text(size = size_title, face = "bold", hjust = 0.5),
    axis.title = element_text(size = size_axis),
    axis.text.x = element_text(size = size_label, color = "black", angle = 45, hjust = 1),
    legend.title = element_text(size = size_axis, face = "bold"),
    legend.text = element_text(size = size_label),
    strip.text = element_text(size = 13, face = "bold"),
    panel.grid = element_blank()
  )

# Display the plot
print(plot_Fun_Phylum + theme(legend.position="right") + guides(fill=guide_legend(ncol= 1)))
```
```{r}
ggsave("D:/Biocrust_Reawakening/iscience/Figures/FigureS_TaxCom_Phylum_Bac.png", 
       plot = last_plot(), 
       device = "png", 
       width = 180,    # 180mm is standard full-page width
       height = 120, 
       units = "mm", 
       dpi = 300)
```

```{r}
ggsave("D:/Biocrust_Reawakening/iscience/Figures/Figure3A_TaxCom_Genus_Bac.pdf", 
       plot = last_plot(), 
       device = "pdf", 
       width = 250,    # 180mm is standard full-page width
       height = 150, 
       units = "mm", 
       dpi = 300)
```

### Beta diversity with NMDS plot analysis

```{r}
sample_data(physeq.LAC)$Incubation = factor(sample_data(physeq.LAC)$Incubation, levels = c("0.05", "1", "24", "48", "96"))

sample_data(physeq.CLC)$Incubation = factor(sample_data(physeq.CLC)$Incubation, levels = c("0.05", "1", "24", "48", "96"))
```

```{r}
physeq_ord <- ordinate(physeq.CLC,method = "NMDS",distance =  "wunifrac")
```

```{r}
ps.dist.all = phyloseq::distance(physeq.CLC, "wunifrac")
```

```{r}
adonis2(ps.dist.all ~Incubation, as(sample_data(physeq.CLC),"data.frame"))
```

```{r}
jpeg("./Baceria/Figures/Bac_Beta_CLC.jpeg", units="in", width = 6.5, height = 4.5, res = 600 )

pslayer = plot_ordination(physeq.CLC, physeq_ord, type = "samples", color = "Incubation") + 
  ggtitle("Bacterial Beta Diversity of Cyanolichen crust") + 
  annotate( "text",
  x = 0.001, y = 0.02, label = "PERMANOVA*','~~italic(p)==0.001*','~~R^2==0.418", parse = TRUE) +
  geom_point(size=4) + 
  scale_colour_manual(values =  c("#F8766D" , "#A3A500" , "#00BF7D",  "#00B0F6", "#E76BF3"),
                      labels = c("0.05H", "1H", "24H", "48H", "96H")) + 
  theme_pubr(legend = "right", border = TRUE) +
  theme(
    text = element_text(family = main_font),
    plot.title = element_text(size = size_title, face = "bold", hjust = 0.5),
    axis.title = element_text(size = size_axis),
    axis.text.x = element_text(size = size_label, hjust = 1),
    legend.title = element_text(size = size_axis, face = "bold"),
    legend.text = element_text(size = size_label),
    strip.text = element_text(size = 13, face = "bold"),
    panel.grid = element_blank()
  ) + 
  theme(plot.title = element_text(hjust = 0.5))

pslayer
```
--------------------------------------------------------------------------------------------------------------------------------
# Alpha Diversity

```{r}
sample_data(physeq.LAC)$Incubation = factor(sample_data(physeq.LAC)$Incubation, levels = c("0.05", "1", "24", "48", "96"))

sample_data(physeq.CLC)$Incubation = factor(sample_data(physeq.CLC)$Incubation, levels = c("0.05", "1", "24", "48", "96"))
```

```{r}
alpha.diversity = estimate_richness(physeq.CLC, measures = c("Shannon"))
data.anova = cbind(sample_data(physeq.CLC), alpha.diversity)
#leveneTest(Observed ~ Incubation, data = data.anova)
physeq.anova = aov(Shannon ~ Incubation, data.anova)
summary(physeq.anova)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.anova, 'Incubation', conf.level = 0.95)

generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }

LABELS=generate_label_df(tukey.ps, 'Incubation')
names(LABELS) = c('Letters','Incubation')
```

```{r}
LABELS
```

```{r}
ylabel <- data.frame("ylabel" = c(6.0, 6.0, 6.0, 6.0, 6.0)) #LAC surface
LABELS$ylabel<-ylabel$ylabel
```

```{r}
jpeg("./Baceria/Figures/Bac_Alpha_CLC_Shannon.jpeg", units="in", width = 5.5, height = 4.5, res = 600 )

#pdf("D:/Biocrust_Reawakening/Github/Fungi(ITS)/Figures/BAC_Alpha_CLC_Shannon.pdf",width = 8.5, height = 5.5)

custom_palette <- c("0.05" = "#F8766D", "1" = "#A3A500", "24" = "#00BF7D", "48" = "#00B0F6", "96" = "#E76BF3")

physeq.plot.richness.incubation <- plot_richness(physeq.CLC, x = "Incubation", color = "Incubation", measures = c("Shannon")) + 
  geom_boxplot(aes(fill = Incubation), color = "black", lwd = 0.7) + 
  ggtitle("Bacterial Alpha Diversity of Cyanolichen crust") + 
  annotate("text", x = 3, y = 6.5, label = "ANOVA, p = 0.003") +
  scale_x_discrete(labels = c("0.05H", "1H", "24H", "48H", "96H")) +
  theme_pubr(border = TRUE, legend = "none") + 
  
  theme(
    text = element_text(family = main_font),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    strip.text = element_text(size = 14, face = "bold"),  # Shannon
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_colour_manual(values = custom_palette) +
  scale_fill_manual(values = custom_palette) +
  geom_text(data = LABELS, aes(x = Incubation, y = ylabel, label = Letters))

physeq.plot.richness.incubation
#dev.off()
```


#Alpha diversity of Phylum
```{r}
physeq.cb = subset_taxa(physeq.LAC, Phylum == "Cyanobacteria")
physeq.cb
```

```{r}
sample_data(physeq.cb)$Incubation = factor(sample_data(physeq.cb)$Incubation, levels = c("0.05", "1", "24", "48", "96"))
```

```{r}
alpha.diversity = estimate_richness(physeq.cb, measures = c("Observed"))
data.anova = cbind(sample_data(physeq.cb), alpha.diversity)
#leveneTest(Observed ~ Incubation, data = data.anova)
physeq.anova = aov(Observed ~ Incubation, data.anova)
summary(physeq.anova)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.anova, 'Incubation', conf.level = 0.95)

generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }

LABELS=generate_label_df(tukey.ps, 'Incubation')
names(LABELS) = c('Letters','Incubation')
```

```{r}
LABELS
```

```{r}
ylabel <- data.frame("ylabel" = c(1800, 1800, 1800, 1600, 1600)) #LAC surface
LABELS$ylabel<-ylabel$ylabel
```

```{r}
tiff("D:/Biocrust_Reawakening/New_data/Bacteria/BAC_Alpha_LAC_Observed_Cyanobacteria.tiff", units="in", width = 4.5, height = 6, res = 600 )

#pdf("D:/Biocrust_Reawakening/Github/Fungi(ITS)/Figures/BAC_Alpha_CLC_Shannon.pdf",width = 8.5, height = 5.5)

physeq.plot.richness.incubation = plot_richness(physeq.cb, x="Incubation", color=("Incubation"), measures=c("Observed")) + 
  geom_boxplot(aes(fill = Incubation), color = "black", lwd = 0.7) + 
  ggtitle("Cyanobacteria in LAC crust type") + 
  #stat_compare_means(method = "t.test", label.y = 300) + 
  annotate("text", x = 3, y = 2000, label = "ANOVA, p = 0.0455") +
  theme_pubr(border= TRUE, legend = c("none")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_hue(l = 58) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(data=LABELS, aes(x=LABELS$Incubation, y = LABELS$ylabel , label = LABELS$Letters))
physeq.plot.richness.incubation

#dev.off()
```