#Taxonomy composition in Genus level

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
library(car)
```

```{r}
otus <- read.table("./Fungi/Data/CrustTemp2ITSasvUN3.otu_table_fix.txt",header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU <- otu_table(otumat, taxa_are_rows = TRUE)
```

```{r}
taxmat.fix <- read.table("./Fungi/Data/tax_fun_fix.txt", header=T, sep="\t" ,row.names=1)
taxmat.fix <- as(as.matrix(taxmat.fix),"matrix")
TAX.fix <- tax_table(taxmat.fix)
```

```{r}
meta <- read.table("./Fungi/Data/CrustTemp2ITSasvUN3.mapping_file_fix.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
sampleData <- sample_data(meta)
```

```{r}
physeqITS <- phyloseq(OTU,TAX.fix,sampleData)
physeqITS
#tax_table(physeqITS)
```

```{r}
physeqITS <- subset_taxa(physeqITS, Kingdom == "Fungi")
physeqITS

physeq.prune = prune_taxa(taxa_sums(physeqITS) > 1, physeqITS)
physeq.prune
```

### Rarefraction curve

Check read counts: any samples that have very low reads should be removed.
[Ref](http://evomics.org/wp-content/uploads/2016/01/phyloseq-Lab-01-Answers.html)

```{r}
readcount = data.table(as(sample_data(physeq.prune), "data.frame"),
                 TotalReads = sample_sums(physeq.prune), 
                 keep.rownames = TRUE)
setnames(readcount, "rn", "SampleID")

#For plotting, use command below.
SeqDepth = ggplot(readcount, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
```

TotalReads of all the samples can be in this table (select only SampleID and TotalReads columns).
In order to check samples with low number of reads, "order()" can be used to sort "TotalReads" column.
In this dataset, N55.Rhizo has very low number of reads, so will will filter this sample out using the next minimum number of reads.
```{r}
readcount = readcount[order(readcount$TotalReads), c("SampleID", "TotalReads")]
```

```{r}
as.data.frame(readcount)
```

Extract OTU table, transpose and convert to data frame
Extract OTU table from phyloseq object as data frame and use vegan to calculate rarefaction curve
```{r}
otu.rare <- otu_table(physeq.prune)
otu.rare <- as.data.frame(t(otu.rare))
sample_names <- rownames(otu.rare)

#vegan rarecurve 
png("D:/Biocrust_Reawakening/Github/Fungi(ITS)/Figures/Rarefraction.png", units="in", width = 8.5, height = 6, res = 600 )
otu.rarecurve <- rarecurve(otu.rare, step = 100, sample = 1000, label = T)
dev.off()
```
```{r}
otu.rare <- otu_table(physeq.prune)
otu.rare <- as.data.frame(t(otu.rare))
sample_names <- rownames(otu.rare)

#vegan rarecurve 
png("D:/Biocrust_Reawakening/Github/Fungi(ITS)/Figures/Rarefraction.png", units="in", width = 9.5, height = 7, res = 600 )
otu.rarecurve <- rarecurve(otu.rare, step = 100, sample = 50000, label = T)
dev.off()
```

### Taxonomic composition

```{r}
physeq.LAC <- subset_samples(physeq.prune, Crusttype=="LAC")
physeq.LAC

physeq.CLC <- subset_samples(physeq.prune, Crusttype=="CLC")
physeq.CLC
```

```{r}
glom.bac.LAC <- speedyseq::tax_glom(physeq.LAC,taxrank = "Genus")
#glom.bac
#tax_table(glom.bac)

glom.bac.CLC <- speedyseq::tax_glom(physeq.CLC,taxrank = "Genus")
#tax_table(glom.bac)
```

```{r}
bac.abund.LAC <- merge_samples(glom.bac.LAC, "Incubation")
sample_data(bac.abund.LAC)$Incubation <- factor(sample_names(bac.abund.LAC))
bac.abund.LAC = transform_sample_counts(bac.abund.LAC, function(x) x / sum(x))
bac.abund.LAC

bac.abund.CLC <- merge_samples(glom.bac.CLC, "Incubation")
sample_data(bac.abund.CLC)$Incubation <- factor(sample_names(bac.abund.CLC))
bac.abund.CLC = transform_sample_counts(bac.abund.CLC, function(x) x / sum(x))
bac.abund.CLC
```

```{r}
data_glom.bac.LAC <- speedyseq::psmelt(bac.abund.LAC)
data_glom.bac.LAC$Genus <- as.character(data_glom.bac.LAC$Genus)

data_glom.bac.CLC <- speedyseq::psmelt(bac.abund.CLC)
data_glom.bac.CLC$Genus <- as.character(data_glom.bac.CLC$Genus)
```

```{r}
data_glom.bac.LAC$Genus[data_glom.bac.LAC$Abundance < 0.005] <- "others"
data_glom.bac.CLC$Genus[data_glom.bac.CLC$Abundance < 0.005] <- "others"
```

```{r}
# Extract percentage data
percentage_data <- data_glom.bac.CLC %>%
  group_by(Incubation, Genus) %>%
  summarize(Percentage = sum(Abundance)) %>%
  group_by(Incubation) %>%
  mutate(Percentage = Percentage / sum(Percentage) * 100) %>%
  gather(key = "Genus", value = "Percentage", -Incubation)

# View the percentage data
print(percentage_data)
```

```{r}
write.csv(percentage_data, file = "D:/Biocrust_Reawakening/New_data/Fungi/CLC_Genus_percentage_data.csv", row.names = FALSE)
```

```{r}
Count.LAC = length(unique(data_glom.bac.LAC$Genus))
Count.LAC

Count.CLC = length(unique(data_glom.bac.CLC$Genus))
Count.CLC
```

```{r}
# Merge the two datasets for consistent color mapping
data_glom_combined <- bind_rows(
  mutate(data_glom.bac.LAC, Group = "LAC"),
  mutate(data_glom.bac.CLC, Group = "CLC")
)
```
```{r}
unique(data_glom_combined$Genus)
```


```{r}
#for LAC Order
data_glom_combined$Genus <- factor(data_glom_combined$Genus, levels = c(
"unclassified_k_Fungi" ,
"unclassified_p_Ascomycota" ,
"unclassified_c_Dothideomycetes"  ,
"unclassified_c_Sordariomycetes", 
"Podospora" ,
"Fimetariella",
"unclassified_o_Pleosporales"  ,  
"unclassified_f_Didymellaceae"  ,
"unclassified_f_Pleosporaceae" , 
"Leptosphaerulina" ,
"Macroventuria" ,
"Nothophoma" ,
"Thyrostroma" ,
"Alternaria" ,
"Preussia" ,
"Deniquelata",
"unclassified_o_Hypocreales"  ,  
"Fusarium" ,
"unclassified_o_Eurotiales" ,   
"unclassified_o_Dothideales" , 
"unclassified_o_Lichenostigmatales",
"unclassified_o_Trapeliales" ,  
"unclassified_o_Botryosphaeriales" ,
"unclassified_o_Onygenales"  ,
"unclassified_o_Chaetothyriales" ,
"Coniochaeta"  ,
"Coniosporium",
"unclassified_o_Helotiales", 
"unclassified_f_Rutstroemiaceae",
"unclassified_o_Orbiliales" , 
"unclassified_o_Pezizales",  
"unclassified_o_Xylariales" ,
"unclassified_o_Capnodiales" ,
"unclassified_o_Umbilicariales",
"unclassified_p_Basidiomycota" ,
"unclassified_o_Tremellales",
"unclassified_o_Agaricales" ,  
"Amanita" ,
"Entoloma",
"Naganishia",
"Filobasidium",
"Basidiobolus"






))
```
```{r}
data_glom_combined$Phylum <- factor(data_glom_combined$Phylum, levels = c(
"Ascomycota" ,
"Basidiomycota" ,
"Basidiobolomycota" ,
"Unknown"
))
```

```{r}
data_glom_combined$Incubation = factor(data_glom_combined$Incubation, levels = c("0.05","1","24","48","96"))
data_glom_combined$Group <- factor(data_glom_combined$Group, levels = c("LAC", "CLC"))
```

```{r}
# Italicize only the genus (first word)
genus_labels <- function(x) {
  sapply(x, function(i) {
    if (grepl("^unclassified", i) || grepl("^[A-Z0-9-]+$", i)) {
      # Leave unclassified or code-only names unchanged
      i
    } else {
      parts <- strsplit(i, " ", fixed = TRUE)[[1]]
      if (length(parts) == 1) {
        paste0("italic('", parts[1], "')")
      } else {
        paste0(
          "italic('", parts[1], "')~'",
          paste(parts[-1], collapse = " "),
          "'"
        )
      }
    }
  })
}
```

```{r}

# The palette with black:
cbbPalette <-
  c(
"unclassified_k_Fungi" = "#CECC9C",
"unclassified_p_Ascomycota" = "tomato4" ,
"unclassified_c_Dothideomycetes" = "indianred2" ,
"unclassified_c_Sordariomycetes" = "#ce4616", 
"Podospora" = "orangered2" ,
"Fimetariella" =  "orangered",
"unclassified_o_Pleosporales" = "darkgreen"  ,  
"unclassified_f_Didymellaceae" = "darkolivegreen4", 
"unclassified_f_Pleosporaceae" = "seagreen1", 
"Leptosphaerulina" =   "darkolivegreen2",
"Macroventuria" = "forestgreen" ,
"Nothophoma" = "seagreen",
"Thyrostroma" = "#00FF00",
"Alternaria" =  "limegreen",
"Preussia" = "seagreen3",
"Deniquelata" = "darkolivegreen", 
"unclassified_o_Hypocreales" = "darkorchid4" ,  
"Fusarium" = "darkorchid1",
"unclassified_o_Eurotiales" = "#5e70de",   
"unclassified_o_Dothideales" = "darkgoldenrod1", 
"unclassified_o_Lichenostigmatales" = "goldenrod3",
"unclassified_o_Trapeliales" = "#f08f6d",  
"unclassified_o_Botryosphaeriales" =  "sienna3",
"unclassified_o_Onygenales" = "antiquewhite4" ,
"unclassified_o_Chaetothyriales" =  "burlywood1",
"Coniochaeta" = "burlywood3" ,
"Coniosporium" = "burlywood2",
"unclassified_o_Helotiales" = "deeppink" , 
"unclassified_f_Rutstroemiaceae" = "deeppink3",
"unclassified_o_Orbiliales" = "#283cb7" , 
"unclassified_o_Pezizales"= "salmon" ,  
"unclassified_o_Xylariales" =   "dodgerblue3" ,
"unclassified_o_Capnodiales" = "indianred",
"unclassified_o_Umbilicariales" = "violetred4",
"unclassified_p_Basidiomycota" = "gold4" ,
"unclassified_o_Tremellales" = "yellow1",
"unclassified_o_Agaricales" = "gold3",  
"Amanita" = "gold",
"Entoloma" = "yellow3",
"Naganishia" = "gold2",
"Filobasidium" = "khaki1",
"Basidiobolus" = "yellow2"
)
```

```{r}
cbbPalette <- c(
  Ascomycota = "#802B51",
  Basidiomycota = "#CDCD42",
  Basidiobolomycota = "#F4BC43",
  Mucoromycota = "#6092C8",
  Blastocladiomycota = "#E29E37",
  Mortierellomycota = "#3A8489",
  Rozellomycota = "#C06EC4",
  Chytridiomycota = "#458933",
  Entomophthoromycota = "#BC3830",
  Unknown = "#CECCA1"
)
```

```{r}
# Ensure consistent levels for the "Genus" factor
data_glom_combined$Genus <- factor(data_glom_combined$Genus, levels = names(cbbPalette))
```

```{r}
# Define your standard sizes
size_title <- 14
size_axis <- 14
size_label <- 10
main_font <- "serif" # Time news roman
```

```{r}
facet_wrap(
  ~Group,
  labeller = labeller(Group = c(
    LAC = "Light cyanobacteria crust",
    CLC = "Cyanolichen crust"
  ))
)
```

```{r fig.height=5, fig.width=7.5, fig.align="center"}
jpeg("./Fungi/Figures/Taxcom_Genus_Fun_0.05.jpeg", units="in", width = 10, height = 7, res = 600 )

# Create the combined plot
plot_Fun_Phylum <- ggplot(
  data_glom_combined,
  aes(x = Incubation, y = Abundance, fill = Genus)
) +
  geom_bar(stat = "identity", position = "fill") +
  scale_x_discrete(labels = c("0.05H", "1H", "24H", "48H", "96H")) +
  scale_fill_manual(values = cbbPalette,
                    labels = function(x) parse(text = genus_labels(x))) +
  facet_wrap(
    ~Group,
    labeller = labeller(Group = c(
      LAC = "Light cyanobacterial crust",
      CLC = "Cyanolichen crust"
    ))
  ) +
  ggtitle("Fungal Taxonomic Composition") +
  theme_bw() +
  theme_pubr(border = TRUE) +
  theme(
    text = element_text(family = main_font),
    plot.title = element_text(size = size_title, face = "bold", hjust = 0.5),
    axis.title = element_text(size = size_axis),
    axis.text.x = element_text(size = size_label, color = "black", angle = 45, hjust = 1),
    legend.title = element_text(size = size_axis, face = "bold"),
    legend.text = element_text(size = size_label),
    strip.text = element_text(size = 13, face = "bold"),
    panel.grid = element_blank()
  )

# Display the plot
print(plot_Fun_Phylum + theme(legend.position="right") + guides(fill=guide_legend(ncol= 2)))
```
```{r}
ggsave("D:/Biocrust_Reawakening/iscience/Figures/Figure3A_TaxCom_Genus_Fun.png", 
       plot = last_plot(), 
       device = "png", 
       width = 250,    # 180mm is standard full-page width
       height = 150, 
       units = "mm", 
       dpi = 300)
```


#### Alpha diversity

```{r}
sample_data(physeq.LAC)$Incubation = factor(sample_data(physeq.LAC)$Incubation, levels = c("0.05", "1", "24", "48", "96"))

sample_data(physeq.CLC)$Incubation = factor(sample_data(physeq.CLC)$Incubation, levels = c("0.05", "1", "24", "48", "96"))
```

```{r}
alpha.diversity = estimate_richness(physeq.CLC, measures = c("Shannon"))
data.anova = cbind(sample_data(physeq.CLC), alpha.diversity)
#leveneTest(Observed ~ Incubation, data = data.anova)
physeq.anova = aov(Shannon ~ Incubation, data.anova)
summary(physeq.anova)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.anova, 'Incubation', conf.level = 0.95)

generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }

LABELS=generate_label_df(tukey.ps, 'Incubation')
names(LABELS) = c('Letters','Incubation')
```

```{r}
LABELS
```

```{r}
ylabel <- data.frame("ylabel" = c(4, 4, 4, 3.5, 3.5)) #LAC surface
LABELS$ylabel<-ylabel$ylabel
```

```{r}
jpeg("./Fungi/Figures/Fun_Alpha_CLC_Shannon.jpeg", units="in", width = 5.5, height = 4.5, res = 600 )

#pdf("D:/Biocrust_Reawakening/Github/Fungi(ITS)/Figures/BAC_Alpha_CLC_Shannon.pdf",width = 8.5, height = 5.5)

custom_palette <- c("0.05" = "#F8766D", "1" = "#A3A500", "24" = "#00BF7D", "48" = "#00B0F6", "96" = "#E76BF3")

physeq.plot.richness.incubation <- plot_richness(physeq.CLC, x = "Incubation", color = "Incubation", measures = c("Shannon")) + 
  geom_boxplot(aes(fill = Incubation), color = "black", lwd = 0.7) + 
  ggtitle("Fungal Alpha Diversity of Cyanolichen crust") + 
  annotate("text", x = 3, y = 4.5, label = "ANOVA, p = 0.0141") +
  scale_x_discrete(labels = c("0.05H", "1H", "24H", "48H", "96H")) +
  theme_pubr(border = TRUE, legend = "none") + 
  
  theme(
    text = element_text(family = main_font),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    strip.text = element_text(size = 14, face = "bold"),  # Shannon
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_colour_manual(values = custom_palette) +
  scale_fill_manual(values = custom_palette) +
  geom_text(data = LABELS, aes(x = Incubation, y = ylabel, label = Letters))

physeq.plot.richness.incubation
```

#Alpha diversity of targeted taxa

```{r}
physeq.cb = subset_taxa(physeq.LAC, Order == "Pleosporales")
physeq.cb
```

```{r}
sample_data(physeq.cb)$Incubation = factor(sample_data(physeq.cb)$Incubation, levels = c("0.05", "1", "24", "48", "96"))
```

```{r}
alpha.diversity = estimate_richness(physeq.cb, measures = c("Observed"))
data.anova = cbind(sample_data(physeq.cb), alpha.diversity)
#leveneTest(Observed ~ Incubation, data = data.anova)
physeq.anova = aov(Observed ~ Incubation, data.anova)
summary(physeq.anova)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.anova, 'Incubation', conf.level = 0.95)

generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }

LABELS=generate_label_df(tukey.ps, 'Incubation')
names(LABELS) = c('Letters','Incubation')
```

```{r}
LABELS
```

```{r}
ylabel <- data.frame("ylabel" = c(80, 80, 60, 80, 80)) 
LABELS$ylabel<-ylabel$ylabel
```

```{r}
tiff("D:/Biocrust_Reawakening/New_data/Fungi/Figures/Fun_Alpha_LAC_Observed_Pleosporales.tiff", units="in", width = 4.5, height = 6, res = 600 )

#pdf("D:/Biocrust_Reawakening/Github/Fungi(ITS)/Figures/BAC_Alpha_CLC_Shannon.pdf",width = 8.5, height = 5.5)

custom_palette <- c("0.05" = "#F8766D", "1" = "#A3A500", "24" = "#00BF7D", "48" = "#00B0F6", "96" = "#E76BF3")

physeq.plot.richness.incubation <- plot_richness(physeq.cb, x = "Incubation", color = "Incubation", measures = c("Observed")) + 
  geom_boxplot(aes(fill = Incubation), color = "black", lwd = 0.7) + 
  ggtitle("Pleosporales in LAC crust type") + 
  annotate("text", x = 3, y = 100, label = "ANOVA, p = 0.00371") +
  theme_pubr(border = TRUE, legend = "none") + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_colour_manual(values = custom_palette) +
  scale_fill_manual(values = custom_palette) +
  geom_text(data = LABELS, aes(x = Incubation, y = ylabel, label = Letters))

physeq.plot.richness.incubation
```

-------------------------------------------------------------------------------------------------------------------------

### Beta diversity with NMDS plot analysis

```{r}
sample_data(physeq.prune)$Incubation = factor(sample_data(physeq.prune)$Incubation, levels = c("0.05", "1", "24", "48", "96"))
sample_data(physeq.prune)$Crusttype = factor(sample_data(physeq.prune)$Crusttype, levels = c("CLC", "LAC"))

sample_data(physeq.LAC)$Incubation = factor(sample_data(physeq.LAC)$Incubation, levels = c("0.05", "1", "24", "48", "96"))

sample_data(physeq.CLC)$Incubation = factor(sample_data(physeq.CLC)$Incubation, levels = c("0.05", "1", "24", "48", "96"))
```

```{r}
physeq_ord <- ordinate(physeq.LAC,method = "NMDS",distance =  "bray")
```

```{r}
ps.dist.all = phyloseq::distance(physeq.LAC, "bray")
```

```{r}
adonis2(ps.dist.all ~Incubation, as(sample_data(physeq.LAC),"data.frame"))
```

```{r}
jpeg("./Fungi/Figures/Fun_Beta_LCC.jpeg", units="in", width = 6.5, height = 4.5, res = 600 )

pslayer = plot_ordination(physeq.LAC, physeq_ord, type = "samples", color = "Incubation") + 
  ggtitle("Fungal Beta Diversity of Light cyanobacterial crust") + 
  annotate( "text",
  x = 1.3, y = 1.2, label = "PERMANOVA*','~~italic(p)==0.001*','~~R^2==0.377", parse = TRUE) +
  geom_point(size=4) + 
  scale_colour_manual(values =  c("#F8766D" , "#A3A500" , "#00BF7D",  "#00B0F6", "#E76BF3"),
                      labels = c("0.05H", "1H", "24H", "48H", "96H")) + 
  theme_pubr(legend = "right", border = TRUE) +
  theme(
    text = element_text(family = main_font),
    plot.title = element_text(size = size_title, face = "bold", hjust = 0.5),
    axis.title = element_text(size = size_axis),
    axis.text.x = element_text(size = size_label, hjust = 1),
    legend.title = element_text(size = size_axis, face = "bold"),
    legend.text = element_text(size = size_label),
    strip.text = element_text(size = 13, face = "bold"),
    panel.grid = element_blank()
  ) + 
  theme(plot.title = element_text(hjust = 0.5))

pslayer
```

```{r}
pslayer = plot_ordination(physeq.prune, physeq_ord, type = "samples", color = "Incubation", shape = "Crusttype") + ggtitle("Fungal Beta Diversity (PCoA) Overall") + 
  annotate("text", x = 1.8, y = 0.8, cex = 4.5, label = "PERMANOVA, p = 0.001") +
  geom_point(size=4) +
  scale_colour_manual(values = c("#F8766D" , "#A3A500" , "#00BF7D",  "#00B0F6", "#E76BF3", "red", "blue")) + 
  theme_pubr(legend = "right", border = TRUE) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_ellipse(aes(color = Crusttype), geom = "path", type = "norm", level = 0.95, 
               linewidth = 0.5, linetype = "dashed", show.legend = TRUE
  )
   

pslayer
```

```{r}
tiff("D:/Biocrust_Reawakening/New_data/Fungi/Figures/Fun_Beta_LAC_NMDS.tiff", units="in", width = 6, height = 4.5, res = 600 )
plot

dev.off()

```
