
Title: PCA analysis
Author: Teeratat Kaewjon
Updated: 8/31/2025
---
output:
  word_document: default
  html_document: default
---
```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
library(car)
library(microViz)
library(ComplexHeatmap)
library(circlize)
```

### PCA Analysis (55 samples)  by using microviz package 
In PCA, the distance between points on the plot doesn't have a direct interpretation, as PCA is not based on distance metrics. The positions of points on the plot are determined by their scores along the principal components, and the distances between points reflect their dissimilarity in the multivariate space.


```{r}
met.matrix <- read.table("Metabolite_matrix_fix.txt",header=T,sep="\t",row.names=1, check.names = FALSE)
metmat <- as(as.matrix(met.matrix), "matrix")
MET = otu_table(metmat, taxa_are_rows = TRUE)
```

```{r}
taxmet <- read.table("Metabolite_information.txt", header=T,sep="\t",row.names=1)
taxmet <- as(as.matrix(taxmet),"matrix")
TAX.met = tax_table(taxmet)
```

```{r}
meta.met = read.table("Metabolomics_metadata.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
sampleData.met <- sample_data(meta.met)
```

```{r}
physeq.met = phyloseq(MET,TAX.met,sampleData.met)
physeq.met
```

```{r}
physeq.met.wo <- subset_samples(physeq.met, Sample_type != "Water")
physeq.met.wo
physeq.met.wo_crl <- microbiome::transform(physeq.met.wo, "clr")
```

###PCA plot w/o water 
```{r}
custom_palette <- c("0.05" = "#F8766D" , "1"="#A3A500", "24"="#00BF7D", "48"= "#00B0F6","96"= "#E76BF3")

# Set the factor levels for the "Samples" variable
sample_data(physeq.met.wo_crl)$Incubation_time <- factor(sample_data(physeq.met.wo_crl)$Incubation_time, levels = c("0.05", "1", "24", "48", "96"))
```

### Save file
```{r}
tiff("D:/Biocrust_Reawakening/Github/Metabolite/Figures/Metabolite_PCA.wo_water.tiff", units="in", width = 8, height = 6, res = 600 )

#pdf("D:/Biocrust_Reawakening/Github/Metabolite/Figures/Metabolite_PCA.wo_water.pdf",width = 8, height = 6)

# Create the PCA plot
physeq.met.wo_crl %>%
  ord_calc(method = "auto") %>%
  ord_plot(color = "Incubation_time", shape = "Sample_type", size = 3.5) +
  scale_colour_manual(values = custom_palette)
```